---
title: "GLIPH Output Analysis"
output:
  html_document:
    df_print: paged
---
```{r, message=FALSE}
library(dplyr)
library(stringr)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(tidyr)
```

Load the data
```{r}
#Load all data
all <- read.table("all.tsv", header = TRUE, sep = '\t')

#Load the group convergence file
d7_group <- read.delim("d7_act_convergence_groups.txt", header = FALSE, sep = '\t')

#Load the motif file
d7_motifs <- read.delim("d7_act.txt", header = TRUE, sep = '\t')
```

```{r}
#Add time column to all data frame
all$time <- paste(all$day, all$status, all$year)

#Re-ordering the level
all$time <- factor(all$time, levels=c("day0 resting 2016", "day0 active 2016", "day7 resting 2016", "day7 active 2016", "day0 resting 2017", "day0 active 2017", "day7 resting 2017", "day7 active 2017"))

#Get the motif name 
motif <- d7_motifs[1]
```

# GLIPH GROUPING
```{r}
#Filter the data 
grp <- d7_group %>%
  filter(V1 > 10) 

#Reorder the data in descending order
grp <- grp[order(-grp$V1),]
```

Get the top 5 groups with highest count  
```{r}
#Get the first 5 rows
grp <- grp[1:5,]

#Separate the aminoAcid
grp <- separate_rows(grp, V3, sep = " ", convert = FALSE)
grp <- as.data.frame(grp)[,-1]
grp$V2 <- as.factor(grp$V2)
```

Get details of the aminoAcids 
```{r}
grp_all <- data.frame()

#Find the aminoAcids that belong the groups
for(i in 1:nrow(grp))
{
  temp <- all[str_detect(all$aminoAcid, grp[i,2]),]
  temp$group <- grp[i,1]
  
  grp_all <- rbind(grp_all, temp)
}

grp_all$group <- as.factor(grp_all$group)
```

Contribution of individuals to the group
```{r}
#Subset d7 active data
dat <- subset(grp_all, time == "day7 active 2016" | time == "day7 active 2017")

#Count the subject based on group 
df <- dat %>% 
  count(patient, group, sort = TRUE)
df
```

```{r, fig.height=3, fig.width=5}
#Control the color of the subject
colorCount <- length(unique(df$patient))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

p <- ggplot(df, aes(x = group, y = n, fill = patient)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = getPalette(colorCount))
p

#To make the plot interactive 
#ggplotly(p, width = 900)
```

Get aminoAcid with count greater than or equal to 9 in either day7 active 2016 or 2017
```{r}
#Sort the df by count 
grp_all <- grp_all[order(-grp_all$count),]

#Get amino acid where the count is greater or equal to 9 in day7 active 
aa <- subset(grp_all, time == "day7 active 2016" & count >=9 | time == "day7 active 2017" & count >=9, aminoAcid)
test <- grp_all[str_detect(grp_all$aminoAcid, paste(aa$aminoAcid, collapse = "|")),]
```

Changes in aminoAcid count across different time point 
```{r, message = FALSE}
df <- test %>%
  group_by(aminoAcid, time) %>%
  summarise(total = sum(count))

df
```

Visualise the data
```{r, fig.width= 5}
p <- ggplot(df, aes(x = time, y = aminoAcid, fill = total, label = total)) + 
  geom_tile() + 
  scale_fill_gradient(low = "dodgerblue4", high = "red3")

p + geom_text()

#To make the plot interactive 
#ggplotly(p, width = 1000)
```

Distribution of aminoAcid across subjects and time points
```{r}
#Group the data by aminoAcid, time, amd patient 
df <- test %>%
  group_by(aminoAcid, time, patient) %>%
  summarise(total = sum(count))

df
```

Visualise the data
```{r}
#Controls the color of the aminoAcid
colorCount <- length(unique(df$aminoAcid))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

p <- ggplot(df, aes(x = patient, y = total, fill = aminoAcid)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = getPalette(colorCount)) +
  facet_wrap(~time, ncol = 2)

p

#To make the plot interactive 
#ggplotly(p)
```

Show that one aminoAcid can have more than one vGeneName 
```{r, message = FALSE}
df <- test %>%
  group_by(aminoAcid, vGeneName, time) %>%
  summarise(total = sum(count))
df
```

Visualise the data
```{r}
colorCount <- length(unique(df$vGeneName))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

p <- ggplot(df, aes(x = aminoAcid, y = total, fill = vGeneName)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = getPalette(colorCount)) +
  facet_wrap(~time, ncol = 2) +
  coord_flip()

p 

#To make the plot interactive 
#ggplotly(p)
```

# GlIPH motif analysis
```{r}
#Subset data for day7 active 2016 and day7 active 2017
d7_act <- subset(all, time == "day7 active 2016" | time == "day7 active 2017")
d7_act_df <- data.frame()

#get the aminoAcids that belong to each individual motif
for(i in 1:nrow(motif))
{
  temp <- d7_act[str_detect(d7_act$aminoAcid, motif[i,]),]
  temp$motif <- motif[i,]
  
  d7_act_df <- rbind(d7_act_df, temp)
}
```

Get the top5 motifs with the highest count 
```{r, message = FALSE}
#Group the data based on motif and time and get the total count
df <- d7_act_df %>%
  group_by(motif, time) %>%
  summarise(total = sum(count))

#re-order the data in descending order by total
df <- df[order(-df$total),]

#Get the top 5 motifs
top5.motif <- head(unique(df$motif), 5)

#Subset the data based on the top5 motif
top5.df <- df[str_detect(df$motif, paste(top5.motif, collapse = "|")),]
top5.df
```

Visualise the top5 motif 
```{r, fig.width = 4, fig.height = 2}
p <- ggplot(top5.df, aes(x = motif, y = time, fill = total, label = total)) + 
  geom_tile() + 
  scale_fill_gradient(low = "dodgerblue4", high = "red3") 

p + geom_text()

##To make the plot interactive 
ggplotly(p)
```

Function to display distribution of aminoAcid for a certain motif across subjects
```{r}
generate_barplot_CDR3 <- function(df, x){
  #Get the aminoAcids that belong to a certain motif 
  df.subset <- subset(df, motif == x)
  
  #Control the color of the aminoAcid
  colorCount <- length(unique(df.subset$aminoAcid))
  getPalette = colorRampPalette(brewer.pal(9, "Set1"))

  p <- ggplot(df.subset, aes(x = patient, y = count, fill = aminoAcid)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = getPalette(colorCount)) +
  ggtitle(paste("CDR3s with", x, "motif")) +
  facet_wrap(~time, ncol = 2)
}
```

Visualise aminoAcid for KTYG
```{r}
p <- generate_barplot_CDR3(d7_act_df, "KTYG")
p

##to make the plot interactive 
#ggplotly(p)
```

Visualise aminoAcid for QSRA
```{r}
p <- generate_barplot_CDR3(d7_act_df, "QSRA")
p

##to make the plot interactive 
#ggplotly(p)
```

Visualise aminoAcid for RANV
```{r}
p <- generate_barplot_CDR3(d7_act_df, "RANV")
p

##to make the plot interactive 
#ggplotly(p)
```
Visualise aminoAcid for SRAN
```{r}
p <- generate_barplot_CDR3(d7_act_df, "SRAN")
p

##to make the plot interactive 
#ggplotly(p)
```

